---
title: "Project 3"
author: "Nfn Tenzin Dakar, Nwe Oo Mon, Crystal Quezada"
date: "2024-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(dplyr)
```

### Data Cleaning and Preparation

The initial step involved cleaning the dataset by removing any unavailable information. We removed the last six rows from the dataset, which contained irrelevant data:

```{r read and clean csv}
joblistings <- read.csv("https://raw.githubusercontent.com/Nweoomon/Project3_607/refs/heads/main/ds_joblistings.csv")
kable(joblistings, format = "markdown", align = 'c')

clean_joblistings <- joblistings %>% 
  slice(1:(n() - 6))
```

### Data Transformation

To facilitate analysis, we transformed the dataset into a long format, organizing it by job source:

```{r transform data}
clean_joblistings_long <- pivot_longer(clean_joblistings, 
                               cols = c("LinkedIn", "Indeed", "SimplyHired", "Monster"), 
                               names_to = "Job_Source", 
                               values_to = "Count")
```

### Job Postings Analysis

The first plot illustrates the total number of job postings across each job source. We calculated the total counts as follows:

```{r job source with most jobs with these skills}
job_counts <- clean_joblistings_long %>%
  mutate(Count = gsub(",", "", Count),
         Count = as.numeric(Count)) %>%
  group_by(Job_Source) %>%
  summarize(Total_Count = sum(Count, na.rm = TRUE), .groups = 'drop')

ggplot(job_counts, aes(x = reorder(Job_Source, Total_Count), y = Total_Count, fill = Job_Source)) +
  geom_bar(stat = "identity") +
  labs(x = "Job Source", y = "Total Job Count", title = "Total Job Counts by Job Source") +
  theme_minimal() +
  geom_text(aes(label = Total_Count), vjust = -0.5, size = 3)
```

### Skill Count Analysis

The next plot provides a comprehensive overview of job listings for various data science skills, sorted by count.

```{r keywords plots}
clean_joblistings_long$Count <- as.numeric(gsub(",", "", clean_joblistings_long$Count))

ggplot(clean_joblistings_long, aes(x = reorder(Keyword, Count), y = Count, fill = Keyword)) +  
  geom_bar(stat = "identity", width = 0.8) +  # Adjust bar width
  labs(title = "Counts of Job Listings by Keyword/Skill",
       x = "Keyword/Skill", y = "Count") +  
  theme_minimal() +  
  coord_flip()
```

### Top Skills Visualization

The subsequent analysis focuses on displaying the top 10 skills by job source:

```{r top 10 skills plot}
top_skills <- clean_joblistings_long %>%
mutate(Count = as.numeric(gsub(",", "", Count))) %>%
  arrange(desc(Count)) %>%
  slice_head(n = 28)

ggplot(top_skills, aes(x = Job_Source, y = Count, fill = Keyword)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = Count), vjust = -1, size = 2.65, position = position_dodge(0.9), size = 3) +
  theme_minimal() +
  labs(title = "Top 10 Data Science Skills by Job Source",
       x = "Job Source",
       y = "Number of Job Listings") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

### Bottom Skills Visualization

Finally, we examine the bottom 10 skills in terms of job listings:

```{r bottom 10 skills plot}
bottom_keywords <- clean_joblistings_long %>%
  mutate(Count = as.numeric(gsub(",", "", Count))) %>% 
  group_by(Keyword, Job_Source) %>%
  summarize(Total_Count = sum(Count), .groups = 'drop') %>% 
  arrange(Total_Count) %>% 
  slice_head(n = 16)

ggplot(bottom_keywords, aes(x = reorder(Keyword, Total_Count), y = Total_Count, fill = Job_Source)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Keyword", y = "Count", title = "Bottom 10 Keywords by Job Source Count") +
  theme_minimal() +
  geom_text(aes(label = Total_Count), vjust = -0.5, position = position_dodge(0.9), size = 3)
```

### Conclusion

